---
title: apply functions
author: Charles T. Gray
date: '2018-07-07'
slug: apply-functions
categories:
  - r-coding
tags:
  - baseR
draft: true
---


```{r load packages, message=FALSE}
# packages 
library(tidyverse)
```


## The good ol' apply functions

I first met this class in my true isolation days. I was on scholarship at a biomedical institute where it felt like nobody thought a second-year university student should be [thus polluting their shades](https://english.stackexchange.com/questions/141532/what-does-this-sentence-mean-heaven-and-earth-are-the-shades-of-pemberley-to).  

I want to know how the function and its arguments can be vecorised.

Should I be experimenting with a list or with a tibble? Let's try a tibble first, hopefully I'm right in thinking of it as a special type of list. One where the cardinality of every element is the same. 

```{r}
# testing tibble
(tt <- tibble(x = c(3, 4), y = c(2, 5)))

```

# Apply a function to each column

```{r}
lapply(tt, function(x) {x + 1})

```

So this seems to work as I understand it. Can I vectorise?

```{r eval=FALSE}
fun_one <- function(x) {x + 1}
fun_two <- function(x) {x^2}

lapply(tt, FUN = c("fun_one", "fun_two"))
```

