---
title: simulations with different distributions
author: Charles T. Gray
date: '2018-07-30'
slug: simulations-with-different-distributions
draft: true
categories:
  - statistics
tags:
  - simulations
  - 3c^2
toc: true
toc_float: true
---


```{r packages, message=FALSE}
# packages

library(tidyverse)

```


### simulation with incomprehensible eval text thing

So I keep hearing about tidy evaluation, and I have a feeling that when I get my heada round, it'll open things up like how tidy did. I thought perhaps I didn't need it, but now I'm back to thinking I do to solve this problem. (Should ask Mike Penguin again.) In the meantime, let's see if I can get a functional simulation, even if I have to eval parse hack text my way through the distribution component.

I have to write up a new raw data simulation, mostly because I can't quite remember the maths of the random sampling for meta-analyses, so I might as well learn how to do it in a purrier way.  

Using this blog post for [inspiration](https://aosmith.rbind.io/2018/06/05/a-closer-look-at-replicate-and-purrr/)^[1], I want to see if I can find a nicer way to writer simulations. (I recently learnt at useR! that I can be described as an incorrigible refactorer.) 

[^1]: Much obliged, [Ariel](https://twitter.com/hashtag/aosmith)!

## meta-parameters

I want to (re)run trials of simulations for:

- different sample sizes
- different distributions
- (also different arm adjustment in the meta-analysis simulation)

```{r}
meta_parameters <- list(
  sample_size = c(15, 30),
  distribution = list(
    list(
      dist = "exp",
      pars = list(rate = 2)
    ),
    list(
      dist = "lnorm",
      pars = list(meanlog = 2, sdlog = 1)
    )
  )
) %>% purrr::cross()

# take a look
meta_parameters %>% str(2)

```


## simulation

```{r}
# simulation function

# nasty eval function - better way?
get_sample <- function(dist, par, n) {
  eval(parse(text =
               paste(
                 c("r",
                   dist,
                   "(",
                   n,
                   ",",
                   paste(par, collapse = ","),
                   ")"),
                 collapse = ""
               )))
}

do_sim <- function(dist = "lnorm",
                   par = c(0, 1),
                   n = 15) {
  
  # is there a better way?
  this_sample <- get_sample(dist, par, n)
  
  # calculate various stats
  this_sample %>% {
    list( # calculate sample statistics
      dist = dist,
      par = as.list(par),
      n = n,
      mean = mean(this_sample),
      median = median(this_sample),
      min = min(this_sample),
      max = max(this_sample),
      q_first = as.numeric(quantile(this_sample, 0.25)),
      q_third = as.numeric(quantile(this_sample, 0.75)),
      iqr = IQR(this_sample)
    ) 
  } 
} %>% Vectorize()

purrr::rerun(3, do_sim())

```

