---
title: "`varameta` <br/> meta-analysis of medians"
subtitle: "or, ruminations on building my first packaged analysis"
author: "Charles T. Gray </br> Fenner School of Environment and Society"
date: "Tuesday 4 September 2018"
output:
  xaringan::moon_reader:
    css: ["default", "rladies", "rladies-fonts"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# in this talk

- quick and dirty meta-analysis
- `varameta::`, a reproducible packaged analyis
- a journey of discovery of tidyverse and related tools

---
class: inverse, bottom
# meta-analysis

---
class: inverse, bottom

# meta-analysis of medians

---
class: inverse, bottom

# `varameta`

---
class: inverse, bottom
# a first packaged analysis

---
class: inverse
#### Thanks for listening! How to bling your slides, r-ladies style: 

```{r eval=FALSE, echo = TRUE}
output:
  xaringan::moon_reader:
    css: ["default", "rladies", "rladies-fonts"]
```

```{r echo = TRUE}
knitr::include_url("https://alison.rbind.io/post/r-ladies-slides/", height = 375)
```


```{bash, echo = FALSE}
# update the online version blog with:
cp -rf ../fenner-2018 ../../../cantabile/static/talks

```

